Grade Retention and Mixed Race

### 2022-11-16 products/paper/main.qmd
Make copy edit corrections

The experiences of mixed race individuals will play an important role in the future of race in the US. A rise in interracial marriage and coupling since the 1960s has resulted in a "biracial baby boom" of individuals whose parents cross racial lines [@root_racially_1992]. This demographic phenomenon may have a potentially transformative effect on racial boundaries [@hochschild_creating_2012; @alba_great_2020]. While racial mixing is nothing new to US history [@morning_who_2000; @gullickson_choosing_2011], this biracial boom is notable for both its demographic size and for a post-Civil Rights environment in which historical norms of racial classification may be eroding. Understanding the transformative potential of this biracial baby boom is contingent on developing a better understanding of how these biracial individuals identify their own race and how they are sorted into a prevailing racial hierarchy that reinforces a system of racial inequality.

Extensive scholarship documents the identification and classification decisions made by or on behalf of multiracial individuals [@xie_racial_1997; @qian_options_2004; @roth_end_2005a; @brunsma_interracial_2005; @herman_forced_2004; @holloway_place_2009; @bratter_mother_2009; @davenport_role_2016; @liebler_boundaries_2016]. This research has generally found a surprising degree of fluidity and choice in classification decisions that is nonetheless somewhat predictable by local context, socioeconomic status and other characteristics. On the other hand, only a smattering of research has documented how the life chances of biracial individuals compare to their monoracial peers [@kao_racial_1999; @campbell_multiracial_2009; @herman_blackwhiteother_2009]. Two important data limitations have contributed to the lack of work on this important topic.

First, using self identification of biracial respondents to analyze outcomes is problematic. Research on the identification and classification of biracial individuals indicates that individuals whose parents belong to different races will not always or consistently identify with both races [@doyle_are_2007; @liebler_america_2017; @gullickson_racial_2019]. Furthermore, social status and social mobility can predict racial identification generally [@saperstein_racial_2012; @saperstein_mulatto_2013] and social status is related to the identification choices of biracial individuals specifically [@roth_end_2005a; @davenport_role_2016]. Thus, differences in outcome among biracial individuals may in part produce differences in identification, rather than the other way around.

This problem of reverse causation can be largely addressed when researchers have access to the racial identification of a respondent's biological parents. While such information is available on some large scale longitudinal surveys, researchers then run into a second problem of sample size. The samples of biracial respondents in such cases are generally far too small to make reliable statistical inferences. This limitation is particularly severe for biracial individuals with two non-white parents.

@@

One important outcome, grade retention, is measurable and important even for young children, however. Being held back in school is a relatively common occurrence that has consequences for later educational transitions [@jimerson_metaanalysis_2001] and is disproportionately applied by race [@warren_patterns_2014]. Grade retention is arguably one of the earliest stratifiers of life chances that an individual will experience and reveals the imprint of the racial hierarchy. In this article, I use data from the American Community Survey to examine the likelihood of being held back in school for biracial children in comparison to their monoracial peers. Pooling data from 2010 to 2019, I am able to draw a large enough sample to make reliable statistical inferences for most biracial groups, including those with two non-white parents.

@@

I distinguish three mechanisms that help us understand where biracials might be positioned within overall system of racial inequality. First, I consider arguments for the racial hierarchy that focus on the effects of racial discrimination at both an individual and systemic level. Second, I consider arguments regarding different distributions of resources across racial groups. Third, I consider arguments regarding how the nature of hybridity itself may affect outcomes.

@@

The strongest historical precedent for understanding such classification in the US is the "one drop rule" norm concerning Black ancestry. According to the one-drop rule, individuals of known Black ancestry are considered exclusively Black, regardless of other ancestries[@davis_who_1991]. @davis_who_1991 generalizes this practice into the concept of *hypodescent* in which individuals with mixed racial ancestry will be classified as members of the lower status group. By comparison a pattern of *hyperdescent* implies that individuals of mixed racial ancestry will be classified as members of the higher status group. This pattern of hyperdescent has historically governed the experiences of individuals with American Indian ancestry [@snipp_american_1989; @wolfe_land_2001; @gullickson_choosing_2011].

Although the hypodescent/hyperdescent paradigm is often used to theoretically frame the possible contemporary experiences of multiracial individuals, there are reasons to be cautious about the generalizability of this paradigm to both the contemporary and historical situation. Hypodescent and hyperdescent are observed in two specific historical cases involving Black ancestry and American Indian ancestry, respectively. As @iverson_regimes_2022 note, these two cases could just as parsimoniously be explained by a "dominance" model which ranks ancestries by their tendencies to be supercessive or recessive in determining identification and classification. In this framework, within the historical "racial triangle" of American colonialism, Black ancestry is supercessive, Indigenous ancestry is recessive, and White ancestry falls somewhere in between.

@@

Status exchange theory predicts that given an unequal racial hierarchy, individuals from a lower status racial group will be positively selected on socioeconomic characteristics into intermarriage with a higher status group, while individuals from a higher status racial group will be similarly negatively selected into the same marriage [@davis_intermarriage_1941; @merton_intermarriage_1941; @fu_racial_2001]. Using education as the socioeconomic parameter, prior research has shown strong evidence of this pattern in the case of Black male/White female marriages in the US, but weaker evidence for other types of unions [@gullickson_education_2006; @fu_racial_2001; @kalmijn_educational_2010; @hou_interracial_2011].

Regardless of its overall prevalence, the overall effects of status exchange on resources are complex because it positively selects one parent and negatively selects the other partner. Therefore the two effects may somewhat cancel each other out in terms of the overall distribution of resources across parents. Complicating this issue further, research has shown that the parents of biracial children may differ from parents of monoracial children in the magnitude and types of resources that they allocate to children [@cheng_constraints_2007], even among parents of similar resources.

@@

Marginal man theory viewed culture in monolithic and static terms and thus to be trapped in between two cultures was seen as a difficult experience. Contemporary views of culture instead emphasize the often flexible, fragmented, and fluid use of cultural bits that comprise a "toolkit" [@swidler_culture_1986; @dimaggio_culture_1997]. From this perspective, biracial individuals may benefit from an enlarged cultural toolkit that allows them to navigate a variety of diverse social contexts [@shih_multiple_2019].

The complexity of the underlying mechanisms described above prevents simple and straightforward hypotheses regarding the position of biracial individuals. Different mechanisms may produce the same observed outcome and the weight of different mechanisms may also vary by groups as these groups may governed by different regimes of classification. However, comparing outcomes across biracial groups may help us better understand the likely mechanisms that drive biracial classification and to disregard others as implausible. Accounting for the effect of parental resources on outcomes to the best of our ability is also key to reaching any conclusions about the placement of biracial individuals within the racial hierarchy.

@@

Results in prior work tend to be imprecise and inconclusive, due to the sample size for most biracial groups. An alternative approach is to use large scale Census or American Community Survey (ACS) data to identify biracial children by the race of the biological parents in their household. This approach can generate much larger samples of biracial individuals, although it does exclude examination of children living in non-nuclear households. Furthermore, researchers then have a limited range of outcomes to explore because biracial individuals identified by this method are by definition children and the data sources include relatively little detailed information. This approach has been used previously to examine the likelihood of living in poverty [@bratter_multiracial_2013; @bratter_poverty_2013] and residential segregation [@ellis_agents_2012]. These studies provide valuable information on biracial childrens' lived experience, but do not directly measure their own outcomes.

However, one important educational outcome is identifiable for children in Census and ACS data - grade retention. Grade retention is partially identifiable in these data sources and provides one of the earliest measurable forms of stratification that individuals face in their lives.

@@

The final analytical sample consists of children aged 5-20 who are currently enrolled in the K-12 school system and live in a household with two parents who are both likely to be biological parents. This necessary sample restriction will bias the family structure of the analytical sample relative to the total population of K-12 schoolchildren. The result of this bias is well-known. Children living in two-parent nuclear households have greater educational success than other children as a result of material and social resources [@biblarz_family_1999]. Therefore, this sample restriction will bias overall estimates of grade retention downward. However, this bias is less problematic for the goal of this study, which is to estimate differences between groups. The sample restriction effectively eliminates differences in family structure across racial groups which may account for some observed differences in grade retention. However, the remaining differences observed here are likely to hold for the full population in the absence of strong racial differences in the effect of family structure on grade retention.

@@

[^6]: Sensitivity analysis showed that this functional form was preferred by the Bayesian information criterion to models with no interaction and a model with interaction terms that treated year as a categorical variable. I also considered a similar interaction between state and year, but this model was not preferred.

```{r}
#| label: fig-year-grade
#| fig-cap: Trends over time and grade in the percent of students clearly behind grade level. Lines are fit to each set of points by grade via lowess smoothing.

acs %>%
  group_by(current_grade, year) %>%
  summarize(below_grade=mean(below_exp_grade)) %>%
  ggplot(aes(x=year, y=below_grade, group=current_grade, color=current_grade))+
  #geom_line()+
  geom_smooth(se=FALSE)+
  geom_point()+
  scale_x_continuous(breaks=c(2010,2014,2018))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="year", y="percent clearly behind grade level for age",
       color="current grade")+
  theme_bw()
```

@@

I begin by showing the relative risk of being CBEG among monoracial respondents across models. @tbl-monoracial reports average marginal effects on the probability of being CBEG for each monoracial minority group in comparison to White students. Understanding these differences helps to clarify the potential for how biracial respondents might be positioned between monoracial groups.

@@

Model 1 only controls for structural factors such as state of residence, student location, grade, and year. These results provide a baseline estimate of the differences across monoracial groups without controlling for differences in cultural and material resources across groups. The results show that Black, Indigenous, and Latino students all have substantially higher probability of being CBEG than White students. Indigenous students have substantially higher risk than all other students, with a probability of being CBEG that is 4.1% percentage points higher than White students. Asian students have the lowest probability of being CBEG, and their risk is slightly lower than Whites.

@@

In total, resource differentials account for a substantial part of the overall racial differences in the risk of being CBEG, but not all of these differences. Black and Indigenous students remain at higher risk than other students. Notably, Asian students rather than White students have the lowest risk of being CBEG in all models. The most notable shift across models is for Latino students. Their higher risk of being CBEG relative to White students is completely accounted for by resource differentials. When comparing hite and Latino students with the same level or resources in Model 4, Latino students actually have slightly lower risk of being CBEG.

Although, the racial differences in risk of BBEG are substantially reduced by controlling for resource differentials, the remaining gaps across all pairwise combinations are still large enough in most cases to sustain the question of where biracial individuals fit within these gaps. Before turning to this question, however, I want to better understand how the resources of biracial individuals compare to their monoracial constituent groups.

@@

@tbl-resources shows the mean values of the resource variables across monoracial and biracial groups.[^7] Because of the number of variables and groups involved, this table can be overwhelming to read. Therefore, I also provide a summation of these resource differentials by racial group by calculating a counterfactual probability of BBEG for each group based on their observed distribution of resource characteristics. This counterfactual probability is calculated by estimating the average predicted probability for each group from the fullest model (Model 4 of \@tbl-monoracial) when all non-resource variables, including race, are held at their mean.[^8]

@@

The results show different risks of being CBEG across biracial groups because of differences in resources. For example, White/Asian students have the lowest counterfactual risk of being CBEG of all groups at 2.4%, owing to their exceptionally high family income and the high college attainment of their parents. At the other end of the spectrum, Indigenous/Latino students have a counterfactual risk of BBEG nearly 2% higher, given their much lower family resources.

@@

Table 4 summarizes the patterns from @tbl-resources across all ten biracial groups. In six of these cases, the biracial group would have outcomes similar or better than the monoracial group with better resources. These results imply that for the majority of these groups, parents in these matches are being positively selected from their constituent groups.

@@

I now turn to the risk of CBEG for biracial respondents. Due to the number of comparisons being made, I split these results into two separate figures. In @fig-black-models, I show the placement of each biracial group involving one black parent in comparison to their two monoracial constituent groups. To allow comparison across models, I compare the results from the baseline model (Model 1 in @tbl-monoracial) and the full model that accounts for all cultural and family resources (Model 4 in @tbl-monoracial).

@@

Controlling for resource variables has some effect on the placement of part-Black biracial respondents.Black/White biracial students initially have a risk of being CBEG slightly higher than and statistically distinguishable from White students. After controlling for cultural and family resources, however, Black/White studentss have a risk of being CBEG that is slightly lower but not statistically distinguishable from White students. This change reflects the relatively low level of parental education and income among Black/White students, which is much closer to Black students than White students, as shown on @tbl-descriptive.

@@

The role of parental resources is harder to determine for Black/Asian and Black/Indigenous students due to the greater uncertainty in these estimates. Controlling for resources moves the point estimate for Black/Asian students from a point estimate close to Asian students to more of a halfway position between the two constituent groups. Black/Indigenous students have risk of being CBEG roughly halfway between their two constituent groups in both models. However, this finding of a halfway position for both groups is highly tentative owing to the wide confidence intervals around both groups.

@@

@fig-other-models shows the results for the remaining biracial groups. For most of these groups, the full models indicate that each biracial group has a risk of being CBEG similar to the risk of the lower monoracial constituent group or has risk higher than this group but less than the halfway point. There are two exceptions to this general pattern. First, the point estimate of the probability of being CBEG for the Indigenous/Latino group is roughly halfway between the two constituent groups, although it is not statistically distinguishable from the lower group. Second, Latino/Asian students are the only case where the point estimate of the risk for the biracial group is higher than either monoracial group. However, the confidence bands suggest that the risk for this group cannot be statistically differentiated from either of the two monoracial constituent groups who have very similar risks. This result only emerges in the model that controls for resources, because Latino/Asian students come from more advantaged households than both Latino and Asian students, on average.

@@

Results for part-Latino students are complicated by the substantial change in the monoracial Latino risk across models. For example, while Latinos have substantially higher risk of being CBEG than Whites in the baseline model, there is only a very small difference favoring Latinos in the full model. Nonetheless, even in this case, we see that White/Latinos have risk closer to those of the lower Latino group, but distinguishable from both groups.

>>>>

The experiences of mixed race individuals will play an important role in the future of race in the US. A rise in interracial marriage and coupling since the 1960s has resulted in a "biracial baby boom" of individuals whose parents cross racial lines [@root_racially_1992] and this demographic surge may have a transformative effect on racial boundaries [@hochschild_creating_2012; @alba_great_2020]. While racial mixing is nothing new to US history [@morning_who_2000; @gullickson_choosing_2011], this biracial boom is notable for its demographic size and has occurred in a post-Civil Rights environment in which historical norms of racial classification may be eroding.

Understanding the transformative potential of this biracial baby boom is contingent on developing a better understanding of both how these biracial individuals identify their own race and how they are sorted into a prevailing system of racial inequality. While extensive scholarship documents the identification and classification decisions made by or on behalf of multiracial individuals [@xie_racial_1997; @qian_options_2004; @roth_end_2005a; @brunsma_interracial_2005; @herman_forced_2004; @holloway_place_2009; @bratter_mother_2009; @davenport_role_2016; @liebler_boundaries_2016], only a smattering of research has documented how the life chances of biracial individuals compare to their monoracial peers [@kao_racial_1999; @campbell_multiracial_2009; @herman_blackwhiteother_2009]. In this article, I expand on this work by examining the risk of grade retention for K-12 students whose parents identify with different races.

By examining this outcome, I am able to overcome two important limitations that have hampered prior work on the life chances of biracial individuals. First, using self identification of biracial respondents to analyze outcomes is problematic. Research on the identification and classification of biracial individuals indicates that individuals whose parents belong to different races do not always consistently identify with both races [@doyle_are_2007; @liebler_america_2017; @gullickson_racial_2019]. Furthermore, social status and social mobility can predict racial identification generally [@saperstein_racial_2012; @saperstein_mulatto_2013] and social status is related to the identification choices of biracial individuals specifically [@roth_end_2005a; @davenport_role_2016]. Thus, differences in outcomes among biracial individuals may in part produce differences in identification, rather than the other way around.

This problem of reverse causation can be largely addressed when researchers have access to the racial identification of a respondent's biological parents. While such information is available on some large scale longitudinal surveys, researchers then run into a second limitation of sample size. The samples of biracial respondents in such data sources are generally far too small to make reliable statistical inferences. This limitation is particularly severe for biracial individuals with two non-white parents.

@@

However, one important outcome, grade retention, is measurable and important even for young children. Being held back in school is a relatively common occurrence that has consequences for later educational transitions [@jimerson_metaanalysis_2001] and is disproportionately applied by race [@warren_patterns_2014]. Grade retention is arguably one of the earliest stratifiers of life chances that an individual will experience. In this article, I use data from the American Community Survey to examine the likelihood of being held back in school for biracial children in comparison to their monoracial peers. Pooling data from 2010 to 2019, I am able to draw a large enough sample to make reliable statistical inferences for most biracial groups, including those with two non-white parents.

@@

I distinguish three mechanisms that help us understand where biracials might be positioned within a system of racial inequality. First, I consider arguments for the racial hierarchy that focus on the effects of racial discrimination at both an individual and systemic level. Second, I consider arguments regarding different distributions of resources across racial groups. Third, I consider arguments regarding how the nature of hybridity itself may affect outcomes.

@@

The strongest historical precedent for understanding such classification in the US is the "one drop rule" norm concerning Black ancestry. According to the one-drop rule, individuals of known Black ancestry are considered exclusively Black, regardless of other ancestries [@davis_who_1991]. @davis_who_1991 generalizes this practice into the concept of *hypodescent* in which individuals with mixed racial ancestry will be classified as members of the lower status group. By comparison a pattern of *hyperdescent* implies that individuals of mixed racial ancestry will be classified as members of the higher status group. This pattern of hyperdescent has historically governed the experiences of individuals with American Indian ancestry [@snipp_american_1989; @wolfe_land_2001; @gullickson_choosing_2011].

Although the hypodescent/hyperdescent paradigm is often used to theoretically frame the possible contemporary experiences of multiracial individuals, there are reasons to be cautious about the generalizability of this paradigm. Hypodescent and hyperdescent are observed in two specific historical cases involving Black ancestry and American Indian ancestry, respectively. As @iverson_regimes_2022 note, these two cases could just as parsimoniously be explained by a "dominance" model which ranks ancestries by their tendencies to be supercessive or recessive in determining identification and classification. In this framework, within the historical "racial triangle" of American colonialism, Black ancestry is supercessive, Indigenous ancestry is recessive, and White ancestry falls somewhere in between.

@@

Status exchange theory predicts that given an unequal racial hierarchy, individuals from a lower status racial group will be positively selected on socioeconomic characteristics into intermarriage with a higher status group, while individuals from a higher status racial group will be similarly negatively selected into the same marriage [@davis_intermarriage_1941; @merton_intermarriage_1941; @fu_racial_2001]. Prior research has shown substantial evidence of this pattern in the case of Black male/White female marriages in the US, but weaker evidence for other types of unions [@gullickson_education_2006; @fu_racial_2001; @kalmijn_educational_2010; @hou_interracial_2011].

Regardless of its overall prevalence, the net effect of status exchange on resources is complex because status exchange positively selects one parent and negatively selects the other partner. Therefore the two effects may somewhat cancel each other out in terms of the overall distribution of resources across parents. Complicating this issue further, research has shown that the parents of biracial children may differ from parents of monoracial children in the magnitude and types of resources that they allocate to children [@cheng_constraints_2007], even among parents of similar resources.

@@

Marginal man theorists viewed culture in monolithic and static terms and thus to be trapped in between two cultures was seen as a difficult experience. Contemporary views of culture instead emphasize the often flexible, fragmented, and fluid use of cultural bits that comprise a "toolkit" [@swidler_culture_1986; @dimaggio_culture_1997]. From this perspective, biracial individuals may benefit from an enlarged cultural toolkit that allows them to navigate a variety of diverse social contexts [@shih_multiple_2019].

The complexity of the underlying mechanisms described above prevents simple and straightforward hypotheses regarding the position of biracial individuals. Different mechanisms may produce the same observed outcome and the weight of different mechanisms can also vary by groups as these groups may be governed by different regimes of classification. However, comparing outcomes across biracial groups may help us better understand the likely mechanisms that drive biracial classification and to disregard others as implausible. Accounting for the effect of parental resources on outcomes to the best of our ability is also key to reaching any conclusions about the placement of biracial individuals within the racial hierarchy.

@@

Results in prior work tend to be imprecise and inconclusive, due to the sample size for most biracial groups. An alternative approach is to use large scale Census or American Community Survey (ACS) data to identify biracial children by the race of the biological parents in their household. This approach can generate much larger samples of biracial individuals, although it does exclude examination of children living in non-nuclear households. Furthermore, researchers then have a limited range of outcomes to explore because biracial individuals identified by this method are by definition children. This approach has been used previously to examine the likelihood of living in poverty [@bratter_multiracial_2013; @bratter_poverty_2013] and residential segregation [@ellis_agents_2012]. These studies provide valuable information on biracial childrens' lived experience, but do not directly measure their own outcomes.

However, one important educational outcome is partially identifiable for children in Census and ACS data - grade retention. I now turn to a closer examination of this particular outcome.

@@

The final analytical sample consists of children aged 5-20 who are currently enrolled in the K-12 school system and live in a household with two parents who are both likely to be biological parents. This necessary sample restriction will bias the family structure of the analytical sample relative to the total population of K-12 schoolchildren. The result of this bias is well-known. Children living in two-parent nuclear households have greater educational success than other children as a result of material and social resources [@biblarz_family_1999]. Therefore, this sample restriction will bias overall estimates of grade retention downward. However, this bias is less problematic for the goal of this study, which is to estimate differences between groups. The sample restriction effectively eliminates differences in family structure across racial groups that may account for some observed differences in grade retention. However, the remaining differences observed here are likely to hold for the full population in the absence of strong racial differences in the effect of family structure on grade retention.

@@

[^6]: Sensitivity analysis showed that this functional form was preferred by the Bayesian information criterion to models with no interaction and a model with interaction terms that treated year as a categorical variable. I also considered a similar interaction between state and year, but this model was not preferred. Model fit statistics for all senstivity analyses are provided in the supplementary materials.

```{r}
#| label: fig-year-grade
#| fig-cap: Trends over time and grade in the percent of students clearly behind expected grade. Lines are fit to each set of points by grade via lowess smoothing.

acs %>%
  group_by(current_grade, year) %>%
  summarize(below_grade=mean(below_exp_grade)) %>%
  ggplot(aes(x=year, y=below_grade, group=current_grade, color=current_grade))+
  #geom_line()+
  geom_smooth(se=FALSE)+
  geom_point()+
  scale_x_continuous(breaks=c(2010,2014,2018))+
  scale_y_continuous(labels=scales::percent)+
  labs(x="year", y="percent clearly behind expected grade",
       color="current grade")+
  theme_bw()
```

@@

I begin by showing the relative risk of being CBEG among monoracial respondents across models. Table 2 reports average marginal effects on the probability of being CBEG for each monoracial minority group in comparison to White students. Understanding these differences helps to clarify the potential for how biracial respondents might be positioned between monoracial groups.

@@

Model 1 only controls for structural factors such as state of residence, student location, grade, and year. These results provide a baseline estimate of the differences across monoracial groups without controlling for differences in cultural and material resources. The results show that Black, Indigenous, and Latino students all have substantially higher probability of being CBEG than White students. Indigenous students have substantially higher risk than all other students, with a probability of being CBEG that is 4.1% percentage points higher than White students. Asian students have the lowest probability of being CBEG, and their risk is slightly lower than White students.

@@

In total, resource differentials account for a substantial part of the overall racial differences in the risk of being CBEG, but not all of these differences. Black and Indigenous students remain at higher risk than other students. Notably, Asian students rather than White students have the lowest risk of being CBEG in all models. The most notable shift across models is for Latino students. Their higher risk of being CBEG relative to White students is completely accounted for by resource differentials. When comparing White and Latino students with the same level or resources in Model 4, Latino students actually have slightly lower risk of being CBEG.

Although, the racial differences in risk of CBEG are substantially reduced by controlling for resource differentials, the remaining gaps across all pairwise combinations are still large enough in most cases to sustain the question of where biracial individuals fit within these gaps. Before turning to this question, however, I want to better understand how the resources of biracial individuals compare to their monoracial constituent groups.

@@

@tbl-resources shows the mean values of the resource variables across monoracial and biracial groups.[^7] Because of the number of variables and groups involved, this table can be overwhelming to read. Therefore, I also provide a summation of these resource differentials by racial group by calculating a counterfactual probability of CBEG for each group based on their observed distribution of resource characteristics. This counterfactual probability is calculated by estimating the average predicted probability for each group from the fullest model (Model 4 of Table 2) when all non-resource variables, including race, are held at their mean.[^8]

@@

The results show different risks of being CBEG across biracial groups because of differences in resources. For example, White/Asian students have the lowest counterfactual risk of being CBEG of all groups at 2.4%, owing to their exceptionally high family income and the high college attainment of their parents. At the other end of the spectrum, Indigenous/Latino students have a counterfactual risk of CBEG nearly 2% higher, given their much lower family resources.

@@

Table 4 summarizes the patterns from @tbl-resources across all ten biracial groups. In six of these cases, the biracial group would have outcomes similar to or better than the monoracial group with better resources. These results imply that for the majority of these groups, parents in these matches are being positively selected from their constituent groups.

@@

I now turn to the risk of CBEG for biracial respondents. Due to the number of comparisons being made, I split these results into two separate figures. In @fig-black-models, I show the placement of each biracial group involving one black parent in comparison to their two monoracial constituent groups. To allow comparison across models, I compare the results from the baseline model (Model 1 in Table 2) and the full model that accounts for all cultural and family resources (Model 4 in Table 2).

@@

Controlling for resource variables has some effect on the placement of part-Black biracial respondents. Black/White biracial students initially have a risk of being CBEG slightly higher than and statistically distinguishable from White students. After controlling for cultural and family resources, however, Black/White studentss have a risk of being CBEG that is slightly lower but not statistically distinguishable from White students. This change reflects the relatively low level of parental education and income among Black/White students, which is much closer to Black students than White students, as shown on @tbl-resources.

@@

The role of parental resources is harder to determine for Black/Asian and Black/Indigenous students due to the greater uncertainty in these estimates. Controlling for resources moves the point estimate for Black/Asian students from a risk similar to Asian students to more of a halfway position between the two constituent groups. Black/Indigenous students have risk of being CBEG roughly halfway between their two constituent groups in both models. However, this finding of a halfway position for both groups is highly tentative owing to wide confidence intervals.

@@

@fig-other-models shows the results for the remaining biracial groups. For most of these groups, the full models indicate that each biracial group has a risk of being CBEG similar to the risk of the lower monoracial constituent group or has risk higher than this group but less than the halfway point. There are two exceptions to this general pattern. First, the point estimate of the probability of being CBEG for the Indigenous/Latino group is roughly halfway between the two constituent groups, although it is not statistically distinguishable from the lower group. Second, Latino/Asian students are the only case where the point estimate of the risk for the biracial group is higher than either monoracial group. However, the confidence bands suggest that the risk for this group cannot be statistically differentiated from either of the two monoracial constituent groups. This result only emerges in the model that controls for resources, because Latino/Asian students come from more advantaged households than both Latino and Asian students, on average.

@@

Results for part-Latino students are complicated by the substantial change in the monoracial Latino risk across models. For example, while Latino students have substantially higher risk of being CBEG than White students in the baseline model, there is only a very small difference favoring Latino students in the full model. Nonetheless, even in this case, we see that White/Latino students have risk closer to those of the lower Latino group, but distinguishable from both groups.
